---
title: "Shepherd_M6_Lab_DSCI609"
author: "Emily Shepherd"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chicago Crime Data: Assault

The data examined was the assault data from Chicago occurring from August 1, 2020 to August 1, 2021.

```{r, include=FALSE}
#Set working Directory
setwd("/Users/burrisfaculty/Desktop/DSCode/DSCI609/Module6_Lab")

#Imports
library(tidyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(plotly)
library(TSstudio)
library(forecast)
library(MLmetrics)
library(dplyr)
library(viridisLite)
library(viridis)
library(padr)
library(rmdformats)
library(recipes)
library(rgdal)
library (raster)
library(rgeos)
library(spdep)
library(OpenStreetMap)

#Import the Data
raw_data = read.csv("Threeyears_18_19_20_21Aug-1.csv", as.is = T)

#Subset the Data Aug1 2020 to Aug 1st 2021
raw_data %>%
  mutate(Date = ymd_hms(Date),
         Primary.Type = as.factor(Primary.Type))%>%
  mutate(Date = floor_date(Date, unit = 'hours')) %>%
  arrange(Date)%>%
  filter(Date >= "2020-08-01" & Date < "2021-08-01") %>%
  write_csv("One_Year.csv")

data <- read.csv("One_Year.csv")
#Check data
#head(data,5)
#tail(data,5)
#Convert text to numbers
data$Latitude = as.numeric(data$Latitude)
data$Longitude = as.numeric(data$Longitude)

#Remove incidents with no lat/long

data = data[!is.na(data$Latitude),]
#Should not have to do longitude
#data = data[!is.na(data$Longitude),]

#Convert to an object
wgs84 = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
crime_spt = SpatialPointsDataFrame(data[,c('Longitude','Latitude')], data, proj4string = wgs84)
#plot(crime_spt)
neighborhoods = readOGR(dsn = '.', layer = "geo_export_b4cc3c31-7809-42e5-a88f-292b52d17cd9", stringsAsFactors = F)
neighborhoods = spTransform(neighborhoods, wgs84)
#plot(neighborhoods, col=terrain.colors(nrow(neighborhoods)))

crime_spt$zip = over(crime_spt, neighborhoods)$zip
crime_spt = crime_spt[!is.na(crime_spt$zip),]

#THE PLOTS BELOW TAKE A LONG TIME....
#plot(crime_spt, col = '#00000010')
#plot(neighborhoods, border = "gray", add = T)

chi_town <- readOGR(dsn = '.', layer = 'cityboundary', stringsAsFactors = F)
chicago_proj <- crs(chi_town)

crime1 = spTransform(crime_spt, chicago_proj)
neighborhoods1 = spTransform(neighborhoods, chicago_proj)

Assault2020 = crime1[crime1$Primary.Type == "ASSAULT",]
```

## Aggregate By Area

From the resulting map, we can see that higher numbers of assault occur in the southern neighborhoods of the city. We can also observe that the number of assaults is lower in the northern parts of the cities. However, from this map we cannot tell whether it is a significant difference among the different regions. Additionally, we cannot see differences within individual neighborhoods.

```{r, include=FALSE}
#Aggregating by Area

names(Assault2020) #gets headers for data
neighborhoods1$ASSAULT2020 = over(neighborhoods1, Assault2020, fn = length)$Case.Number
```

```{r, echo=FALSE}
spplot(neighborhoods1, zcol = "ASSAULT2020", main = list(label = "2020-21 Assault Crime by Neighborhood in Chicago"))

```

```{r, include = FALSE}
pixelsize = 200
box = round(extent(neighborhoods1)/pixelsize) * pixelsize
template = raster(box, crs = chicago_proj,
                  nrow = (box@ymax - box@ymin)/pixelsize,
                  ncols = (box@xmax - box@xmin)/pixelsize)
Assault2020$PRESENT = 1
raster2020 = rasterize(Assault2020, template, field = "PRESENT", fun = sum)
```

```{r, include=FALSE}
plot(raster2020, xlim= c(1091131,1205199), ylim = c(1813892, 1951669))
plot(neighborhoods1, border ='00000040', add = T)
```

## Heat Map with Gaussian Smoothing

Next, we use Gaussian smoothing methods to produce a heat map that will show clusters of assaults happening in Chicago. Once again, this does not show whether the differences are significant. We are only seeing where there are higher concentrations of assaults. When comparing the heat map to the map of Chicago aggregated by area, we can see that the regions that are higher concentrations of red in the heat map correspond to the pink, orange, and yellow areas of the previous map. A benefit of the heat map is that it gives us a little more detail because the boundaries are not restricted to the neighborhoods.

```{r, include=FALSE}
kernel = focalWeight(raster2020, d = 264, type = 'Gauss')
heat2020 = focal(raster2020, kernel, fun = sum, na.rm = T)
```

```{r, echo=FALSE}
plot(heat2020, xlim = c(1091131, 1205199), ylim = c(1713892, 1951669))
plot(neighborhoods1, border = '#00000040', add = T)

```

## Street Map with GettisGI\* Overlay

In this map, the coloring will be based on z-scores. We can see from the map where the number of assaults are significantly low, cold spots, and also, regions where the number of assaults are significantly high, hot spots. This map shows much more specificity about where assaults have occurred more than expected and likewise, which areas of the city would be considered the safest.

Additionally, we can see some hotspots that appeared "safe" in the original map aggregated by area. One area that particularly stood out was the red spots along Lake Michigan. Conversely, there are several cold spots occurring in areas that showed high levels of assault in the orginal map aggregated by area.

```{r, include=FALSE}
pixelsize = 500
box = round(extent(neighborhoods1)/pixelsize)*pixelsize
template = raster(box, crs = chicago_proj, 
                  nrows = (box@ymax - box@ymin)/pixelsize,
                  ncols = (box@xmax - box@xmin)/pixelsize)
Assault2020$PRESENT = 1
getisraster = rasterize(Assault2020, template, field = "PRESENT", fun = sum)
getisgrid = rasterToPolygons(getisraster)

# Create the list of neighbors

neighbors = poly2nb(getisgrid) #construct neighbor list from polygon list
weighted_neighbors = nb2listw(neighbors, zero.policy = T) #Spatial weights for neighbors lists

#Perform Getis-Ord GI*
getisgrid$HOTSPOT = as.vector(localG(getisgrid$layer, weighted_neighbors))

#GRID IS COLORED ACCORDING TO Z-SCORES
min(getisgrid$HOTSPOT, na.rm = TRUE)
max(getisgrid$HOTSPOT, na.rm = TRUE)
breaks = c(-15, -1.06, -1, 1, 1.96, 15)
palette = c('purple', 'blue', 'green', 'orange', 'red') 
col = palette[cut(getisgrid$HOTSPOT, breaks)]

```

```{r, include=FALSE}
#Make the plot
plot(getisgrid, col = col, border = NA, xlim = c(1091131, 1205199), ylim = c(1813892, 1951669))
plot(neighborhoods1, border = 'gray', add = T)

```

```{r, include = FALSE}
extent(neighborhoods) #Check boundaries
basemap = openmap(upperLeft = c(42.02304, -87.94011), lowerRight = c(41.64454, -87.52414 ), type = "osm")
basemap = openproj(basemap, chicago_proj)
```

```{r, echo=FALSE}
plot(basemap)
plot(getisgrid, col = col, border = NA, add = T)
legend(x = 'bottomleft', inset = 0.05, bg = 'white', col = palette, pch =15,
       legend = c('Cold Spot > 95% Confidence', 'Cold Spot 68 - 95% Confidence', 
                  'Not Significant', 'Hot Spot 68 - 95% Confidence',
                  'Hot Spot > 95% Confidence'))
```
