---
title: "Week 2 Lab"
author: "Emily Shepherd"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/burrisfaculty/Desktop/DSCode/DSCI609")
library(stargazer)
```

The data we are exploring consists of twelve variables related to 14 countries that have been democracies since World War II

1.  Percent Reduction of Gini Coefficient(Gini_Reduction): The average percent reduction of the gini coefficient is reported for each county. The gini coefficient is a measure of inequality. It varies from 0 to 1. The value is 0 when income is evenly distributed. The value is 1 when all income is held by the top decile.

2.  Pretax and Transfer Inequality(Wage_Ineq): This is measured as the proportion of an earner in the 90th percentile of the earners with median income.

3.  Partinsanship: The measure of partinsanship varies from 0 to 1, with left being 0 and right being 1. A perfectly balanced government would have a measure of 0.5.

4.  Voter Turnout: The percent of eligible citizens who actually vote in elections.

5.  Unionization: This is a measure of union density within the country.

6.  Constitutional Veto Points (Veto_Points):This is a composite measure of availability of a veto option, based on the structure of government.

7.  Electoral System: Each country's government was categorized as majoritarian or proportional. A country with a majoritarian system had a value of 0 and proportional had a 1.

8.  Left Fragmentation (Left_Frag): This is calculated by one minus the sum of the squared seat shares of left-wing parties.

9.  Right Overrepresentation (Right_Over): The difference between the legistlative seat share of right-wing parties and thier share of the vote.

10. Real per capita Income (Inc_per_Capita): This is the income per capita in 1985 dollars. 

11. Female Labor Force Participation (Fem_Labor):This is the percentage of female workforce participants in the working age population.

12. Unemployment: This is the unemployment rate.

```{r, include = FALSE}
#Create Vectors
Country <- c("Australia", "Austria", "Belgium", "Canada",
            "Denmark", "Finland", "France", "Germany",
            "Ireland", "Italy", "Japan", "Netherlands",
            "New Zealand", "Norway", "Sweden", "U.K.", "U.S.")
#Percent Reduction of Gini Coefficient(Gini_Reduction): The average percent reduction of the gini coefficient is reported for each county. The gini coefficient is a measure of inequality. It varies from 0 to 1. The value is 0 when income is evenly distributed. The value is 1 when all income is held by the top decile.
Gini_Reduction <- c(23.97, NA, 35.56, 21.26, 37.89, 35.17,
                     25.36, 18.70, NA, 12.13, NA, 30.59, NA, 27.52, 37.89, 22.67, 17.60)
#Pretax and Transfer Inequality(Wage_Ineq): This is measured as the proportion of an earner in the 90th percentile of the earners with median income.
Wage_Ineq <- c(1.7, NA, 1.64, 1.82, 1.58, 1.68, 1.94, 1.70,
               NA,1.63, NA, 1.64, NA, 1.5, 1.58, 1.78,2.07)

#Partinsanship: The measure of partinsanship varies from 0 to 1, with left being 0 and right being 1. A perfectly balanced government would have a measure of 0.5.
Partinsanship <- c(0.47,0.30, 0.36,0.36, 0.35, 0.30, 0.40, 0.39, 0.42, 0.37,
                   0.78, 0.31, 0.43, 0.15, 0.17, 0.52, 0.40)

#Voter Turnout: The percent of eligible citizens who actually vote in elections.
Voter_Turnout <- c(84, 87, 88, 68, 84, 79, 66, 81, 75, 93, 71, 85, 85, 80, 84, 76, 56)

#Unionization: This is a measure of union density within the country.
Unionization <- c(46, 54, 48, 30, 67, 53, 18, 34, 48, 34, 31, 33, 23, 54, 67, 42, 23)

#Constitutional Veto Points (Veto_Points):This is a composite measure of availability of a veto option, based on the structure of government.
Veto_Points <- c(3, 1, 1, 2, 0, 1, 1, 4, 0, 1, 1, 1, 0, 0, 0, 0, 5)

#Electoral System: Each country's government was categorized as majoritarian or proportional. A country with a majoritarian system had a value of 0 and proportional had a 1.
Electoral_System <-c(0, 1, 1, 0, 1, 1, 0, 1, 0, 1, 0, 1, 0, 1,1,0,0)

#Left Fragmentation (Left_Frag): This is calculated by one minus the sum of the squared seat shares of left-wing parties.
Left_Frag <- c(-0.39, -0.18, -0.34, 0.18, -0.40, -0.18, 0.10, -0.13, -0.33,
               0.20, 0.22, 0.18, -0.40, -0.02, -0.40, 0.08, 0)
#Right Overrepresentation (Right_Over): The difference between the legistlative seat share of right-wing parties and thier share of the vote.
Right_Over <- c(0.10, 0.04,0.27, -0.11, 0.07, 0.09, 0.09, 0.15, 0.70,
                0.08, 0.28, -0.36, 0.98, -0.32, -0.03, 0.07, -0.17)

# Real per capita Income (Inc_per_Capita): This is the income per capita in 1985 dollars.
Inc_Per_Capita <- c(109099, 8311, 8949, 11670, 9982, 8661, 9485, 9729,
                    5807, 7777, 7918, 9269, NA, 9863, 9982, 9282, 13651)

# Female Labor Force Participation (Fem_Labor):This is the percentage of female workforce participants in the working age population.
Fem_Labor <- c(46, 51, 43, 48, 63, 66, 51, 51, 37, 38, 56, 35, 47, 52, 63, 54, 53)

#Unemployment: This is the unemployment rate.
Unemployment <- c(4.63, 2.76, 7.89, 6.91, 6.83, 4.48, 4.57, 4.86, 9.09,
                  8.12, 1.77, 4.62, NA, 2.28, 6.83, 5.01, 5.74)
mydata <- data.frame(cbind(Country, Gini_Reduction, Wage_Ineq, Partinsanship, Voter_Turnout,
      Unionization, Veto_Points, Electoral_System, Left_Frag, Right_Over,
      Inc_Per_Capita, Fem_Labor, Unemployment))
#cast characters to doubles
mydata$Fem_Labor <- as.double(mydata$Fem_Labor)
mydata$Gini_Reduction <- as.double(mydata$Gini_Reduction)
mydata$Wage_Ineq <- as.double(mydata$Wage_Ineq)
mydata$Partinsanship <- as.double(mydata$Partinsanship)
mydata$Voter_Turnout <- as.double(mydata$Voter_Turnout)
mydata$Unionization <- as.double(mydata$Unionization)
mydata$Veto_Points <- as.double(mydata$Veto_Points)
mydata$Left_Frag <- as.double(mydata$Left_Frag)
mydata$Right_Over <- as.double(mydata$Right_Over)
mydata$Inc_Per_Capita <- as.double(mydata$Inc_Per_Capita)
mydata$Unemployment <- as.double(mydata$Unemployment)
```

Below you will find the data frame for all the variables:

```{r}
mydata
```
Here are some summary statistics of Unemployment:

```{r}
mean(na.omit(mydata$Unemployment))
median(na.omit(mydata$Unemployment))
max(na.omit(mydata$Unemployment))
```
Here are some summary statistics of Real Income Per Capita:

```{r}

mean(na.omit(mydata$Inc_Per_Capita))
median(na.omit(mydata$Inc_Per_Capita))
max(na.omit(mydata$Inc_Per_Capita))
```

Below you will find a table of summary statistics for all twelve variables:

```{r}
stargazer(mydata, type = "text", title = "Descriptive Statistics", digits = 1, median = TRUE, out = "table.txt")
```